Open source landscape ecology tools

### 2020-07-21 paper/paper.Rmd
adds spatial data processing section

<!-- the role of data preparation - cropping, reprojecting, changing resolution/simplifying, merging/joining -->

<!--MH: Short paragraph about scale; *grainchanger*; DOI: 10.1111/2041-210X.13177 -->

<!-- MH: landscapetools; DOI: 10.1111/2041-210X.13076 -->

<!-- JN: I am also thinking (unsure yet) about mentioning R as a bridge between other software (e.g. GRASSGIS) -->
<!-- rgee -->
<!-- MH: Maybe one short sentence that R allows to connect to other open-source software. We already introduce the topic previously anyhow -->

>>>>

As we mentioned in the previous sections, gathered or obtained data can be represented by different data types and stored in various file formats.
<!-- and read into R using one of the packages from Section \@ref(sec:spatrep). -->
However, several potential obstacles to applying this data for ecological analysis still exist.

The main issue for many landscape ecology studies is related to projections of the used data.
It is not only required to have all of the used datasets in the same projection, but also selection of a proper projection is crucial.
Coordinates in spatial data can represent one of many coordinate reference systems.
Two main groups of coordinate reference systems (CRS) exist, namely geographical and projected, each having many members.
In the first one, geographical, positions are specified by latitude and longitude coordinates in degrees.
<!---->
However, in most landscape ecology studies, projected CRSs, which use some measurement units (e.g., meters), should be utilized.
<!-- Projected CRS  -->
Selected projected CRS should be based on projection that keeps important properties of spatial data intact (e.g., does not distort areas, shapes, distances, angles) and be appropriate for a given study area.
<!-- Where to find advice on how to select projections-->
Gladly, all of the packages from Section \@ref(sec:spatrep) have interfaces for coordinates transformations.

Second possible issue with spatial data is when the obtained data extends over a larger area than the study region.
Then, some spatial preprecessing, such as vector clipping or raster cropping, is needed.
Packages from Section \@ref(sec:spatrep) also make these operations possible.
Additionally, they offer many other operations, such as merging or joining of spatial data, extraction of values from one dataset into another, raster resolution changes, or vector data simplifications.

<!--MH: Short paragraph about scale; *grainchanger*; DOI: 10.1111/2041-210X.13177 -->
<!-- MH: landscapetools; DOI: 10.1111/2041-210X.13076 -->

Finally, there is a number of tools for landscape ecology implemented in some GIS software, such as ... for GRASS GIS or ... for ... .
Gladly, it is possible to control several GIS software directly from R using dedicated packages, such as *rgrass7* for ..., ... for ..., and *rgee* for Google Earth Engine.
